---
title: "Mono Lake"
description: |
  Analysis of Mono Lake levels by year.
output: 
  distill::distill_article:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# Setup Chunk
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(here)
library(tidyverse)
library(janitor)
library(paletteer)
library(sf)
library(gghighlight)

```

```{r}
# Read in the Data and Skip the first 4 rows of the data set 
mono_lake_data <- read_csv(here("data", "mono_lake_annual_levels.csv"), skip = 5) %>% 
clean_names()
```


```{r}
# Create a beautiful graph
ggplot(data = mono_lake_data, aes(x = year, y = lake_level_feet_above_sea_level)) + 
geom_line(color = "aquamarine", size = 1.3)+
theme_classic() +
  
# Set the range for years and lake levels 
scale_x_continuous(breaks = seq(from = 1850, to = 2017, by = 20))+
scale_y_continuous(breaks = seq(from = 6200, to = 6500, by = 10)) + 

# Add lake level events 
geom_hline(yintercept = 6377, color = "brown", linetype = "twodash")+
annotate("text", x = 1888, y = 6379, label = "Land bridges for wildlife emerge (6,377 ft)", color = "brown")+
geom_hline(yintercept = 6392, color = "darkgreen", linetype = "twodash")+
annotate("text", x = 1865, y = 6394, label = "Stable level (6,392 ft)", color = "darkgreen")+
geom_hline(yintercept = 6350, color = "purple", linetype = "twodash")+
annotate("text", x = 1876, y = 6352, label = "Salinity above 150 g/L (6,350 ft)", color = "purple")+
  
# Customize color and size of title and labels 
theme(plot.title = element_text(color = "#0099f9", size = 16, face = "bold", hjust = 0.5), 
  axis.title.x = element_text(color = "#0099f9", size = 14, face = "italic"),
  axis.title.y = element_text(color = "#0099f9", size = 14, face = "italic"))+
  
# Assigning names for key 
  labs(x = "Year", y = "Lake Level (ft)", title = "Lake Levels Above Sea Level by Year")+
  
# State Water Board Annotation
annotate("segment", x = 1999, xend = 2004, y = 6384, yend = 6426) +
annotate(geom = "text", x = 1996, y = 6428, label = "State Water Board Decision (1994)",
           size = 3, hjust = "center-align", color = "blue")+

# Mono Lake Reserve Annotation 
annotate("segment", x = 1981, xend = 1930, y = 6373, yend = 6387) +
annotate(geom = "text", x = 1920, y = 6389, label = "Mono Lake Reserve Created (1981)",
           size = 3, hjust = "center", color = "blue") +

# Restoration Orders Annotation 
annotate("segment", x = 1998, xend = 1986, y = 6383.3, yend = 6416) +
annotate(geom = "text", x = 1974, y = 6419, label = "Restoration Orders Issued (1998)",
           size = 3, hjust = "center", color = "blue")+

# First Study of Mono Lake Annotation 
annotate("segment", x = 1976, xend = 1925, y = 6378.3, yend = 6409) +
annotate(geom = "text", x = 1920, y = 6411, label = "First Study of Mono Lake (1976)",
           size = 3, hjust = "center", color = "blue")
```

F.1 This graph shows the Mono lake levels above sea level (ft) by year.  The key historic events are included as well as the changes in the ecosystem due to changing lake levels. The purple dashed line represents the decrease in brine shrimp, which is the main food supply for various species of birds that rely on them (Eared Grebes and California gulls). The brown line represents the emergence of land bridges that allows predators such as coyotes, to access major bird nesting sites. The green line which represents the lake level stability, shows that the lake levels have become unstable in the recent years since 1975.